% !TEX root =  ../thesis.tex
\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
I would like to thank my supervisor Dr. Stefanos Zafeiriou for being a
remarkable mentor and an inspiring teacher throughout my Ph.D. He has always
provided me unique guidance and showed me how to reach my full potential, while
patiently transferring me his knowledge and way of thinking. I am also thankful
to Prof. Maja Pantic for accepting me in the iBUG group and supporting me with
whatever I needed. I would also like to thank Dr. Georgios Tzimiropoulos for
our collaboration in the early stages of my Ph.D.

Moreover, I would like to express my most genuine gratitude to my friends and
members of the Menpo team $-$ Joan Alabort-i-Medina, James Booth, Patrick Snape
and George Trigeorgis. No words can express the amount of things they all
individually taught me and my Ph.D. experience would never be the same without
our constructive collaboration. My gratitude and appreciation
also goes to all my colleagues at the iBUG group, and especially to Christos
Georgakis, Stefanos Eleftheriadis, Simos Nikitidis and Thanos Papaioannou,
since their friendship and support made these last years an amazing and
unforgettable experience.

Finally, I would like to express my utmost gratefulness to my parents, Petros
and Eleni, and my brother, George, for their continuous and unconditional 
support. Last but foremost, my deepest gratitude goes to my life partner and
soul-mate, Leda, for her boundless love. This thesis is dedicated to them.
\end{abstract}
